---
title: "Assignment1"
subtitle: "SDGB 7844, Prof. Nagaraja"
author: "Yuwen (Suyi) Wu, 330-530PM"
date: "9/22/2018"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)    
library(ggplot2)
```

## Question 1.a


Solution: 

```{r}
class(state.area)
class(state.name)
class(state.region)
```



## Question 1.b What is the longest state name (including spaces)? How long is it?


Solution: 

```{r}
state.name[nchar(state.name)==max(nchar(state.name))]
nchar(state.name[nchar(state.name)==max(nchar(state.name))])
```



## Question 1.c

Compute the average area of the states which contain the word "New" at the start of the state name. Use the function substr().

Solution: 

```{r}
mean(state.area[substr(state.name, start=1, stop=3)=="New"]) 
```




## Question 1.d

Use the function table() to determine how many states are in each region. Use the function kable() to include the table in your solutions. (Notes: you will need the R package knitr to be able to use kable(). See the RMarkdown example in the Assignments folder on Blackboard for an example.)

Solution: 

```{r}

z<-table(state.region)
kable(z)

```




## Question 2.a



Solution: 

```{r}

num.perfect<-2
count<-1 ## change the begin number to get first two perfect number rather than first three perfect number
iter<-2


while(count<=num.perfect){
  divisor<-1
  for(i in 2:iter){ ## when i in 2:iter , i could be the divisor of iter
       if(iter%%i==0 & iter!=i) ## the i could not be the iter in order to get correct sum of i 
         divisor<-c(divisor,i)
         i<-i+1## add this line to iterrate add the i , and make this loop could be finished 
       }# end for loop
  
       if(sum(divisor)==iter){
    print(paste(iter,"is a perfect number",sep=" ")) ## leave a blank space between number and string
    count<-count+1
  }#end if
  
  iter<-iter+1
}#end while loop

```


## Question 2.b.1



Solution: 

```{r}

date()
num.perfect<-2
count<-1 ## change the begin number to get first two perfect number rather than first three perfect number
iter<-2


while(count<=num.perfect){
  divisor<-1
  for(i in 2:iter){ ## when i in 2:iter , i could be the divisor of iter
       if(iter%%i==0 & iter!=i) ## the i could not be the iter in order to get correct sum of i 
         divisor<-c(divisor,i)
         i<-i+1## add this line to iterrate add the i , and make this loop could be finished 
       }# end for loop
  
       if(sum(divisor)==iter){
    print(paste(iter,"is a perfect number",sep=" ")) ## leave a blank space between number and string
    count<-count+1
  }#end if
  
  iter<-iter+1
}#end while loop
date()
```

1 sec


## Question 2.b.2

Solution: 

```{r}

date()
num.perfect<-4
count<-1 ## change the begin number to get first two perfect number rather than first three perfect number
iter<-2


while(count<=num.perfect){
  divisor<-1
  for(i in 2:iter){ ## when i in 2:iter , i could be the divisor of iter
       if(iter%%i==0 & iter!=i) ## the i could not be the iter in order to get correct sum of i 
         divisor<-c(divisor,i)
         i<-i+1## add this line to iterrate add the i , and make this loop could be finished 
       }# end for loop
  
       if(sum(divisor)==iter){
    print(paste(iter,"is a perfect number",sep=" ")) ## leave a blank space between number and string
    count<-count+1
  }#end if
  
  iter<-iter+1
}#end while loop
date()


v<-1:4
time.sec<-c(1,1,1,37)
time.sec
runTime<-cbind(v,time.sec)
runTime

colnames(runTime) <- c("No.perfect number", "waiting time (s)")
runTime
kable(runTime, caption="Run Time Result")
```

[1] "Sat Sep 15 00:27:16 2018"
[1] "6 is a perfect number"
[1] "28 is a perfect number"
[1] "496 is a perfect number"
[1] "8128 is a perfect number"
[1] "Sat Sep 15 00:28:04 2018"



## Question 2.c

 

Solution: 

```{r,fig.width=8, fig.height=6}
theme.info <- theme(plot.title = element_text(size=30, face = "bold", hjust=0.5),
                    axis.title = element_text(size=12),
                    axis.text=element_text(size=20, face = "bold")) 
x<-1:4
y<-c(1,1,1,37)
plot(x,y,pch=20,type = "b",
     xlab = "number of perfect numbers discovered",
     ylab = "cumulative length of time (in seconds)",
     main = "Cumulative Run Times to Discover Perfect Numbers",
     las=TRUE)+theme.info
     
```

Not Linear, it seems exponent function.

## Question 3.a

 
Solution: 

```{r}

x<-c(4,67,3)
x<-x[!is.na(x)] 
sum.x<-1
x
for (i in 1:length(x)){
  if(x[i]<=0){
    print('error')
    sum.x<-NaN
    break
    
  }
  else{
    sum.x<-sum.x*x[i]
    
  }
  }
geometric.mean<-sum.x^(1/length(x)) 
geometric.mean


```



## Question 3.b.1

 

Solution: 

```{r}

x<-c(NA,4,67,3)
x<-x[!is.na(x)] 
sum.x<-1
x
for (i in 1:length(x)){
  if(x[i]<=0){
    print('error')
    sum.x<-NaN
    break
    
  }
  else{
    sum.x<-sum.x*x[i]
    
  }
  }
geometric.mean<-sum.x^(1/length(x)) 
geometric.mean



```

## Question 3.b.2

 

Solution: 

```{r}

x<-c(0,NA,6)
x<-x[!is.na(x)] 
sum.x<-1
x
for (i in 1:length(x)){
  if(x[i]<=0){
    print('error')
    sum.x<-NaN
    break
    
  }
  else{
    sum.x<-sum.x*x[i]
    
  }
  }
geometric.mean<-sum.x^(1/length(x)) 
geometric.mean


```


## Question 3.b.3

 

Solution: 

```{r}

x<-c(67,3,Inf)
x<-x[!is.na(x)] 
sum.x<-1
x
for (i in 1:length(x)){
  if(x[i]<=0){
    print('error')
    sum.x<-NaN
    break
    
  }
  else{
    sum.x<-sum.x*x[i]
    
  }
  }
geometric.mean<-sum.x^(1/length(x)) 
geometric.mean


```

## Question 3.b.4

```{r}

x<-c(67,3,-Inf)
x<-x[!is.na(x)] 
sum.x<-1
x
for (i in 1:length(x)){
  if(x[i]<=0){
    print('error')
    sum.x<-NaN
    break
    
  }
  else{
    sum.x<-sum.x*x[i]
    
  }
  }
geometric.mean<-sum.x^(1/length(x)) 
geometric.mean


```